The Calculus of Linear Constructions

### 2021-12-02 writing/draft5/draft5.tex
draft5

@@

\section{Future Work}
The integration of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.

On the theoretical side, we intend to investigate the semantic models of CLC. In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy. This has since been removed due to the additional difficulty of working with $Prop$ in the soundness proof of CILC. We believe that these difficulties are not intrinsic and $Prop$ could be added back with careful proof work.

>>>>

\subsection{Mutable Data}
Figure \ref{list} presents a CPS append function for linear lists. In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}. Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$. Each \texttt{_cons_} is applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place. Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once. Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.

\begin{figure}[h]
  \caption{Linear Lists}
  \begin{minted}{coq}
  Inductive list (A : U) : L :=
  | nil : list A
  | cons : A -> list A -> list A.

  Fixpoint append (A : U) : list A -o 
    list A -o (list A -o list A) -o list A 
  :=
    fun ls1 ls2 k =>
      match ls1 with
      | nil => k ls2
      | (cons h t) as _cons_ =>
        append _ t ls2 
          (fun res => k (_cons_ h res))
      end.
  \end{minted}
  \label{list}
  \Description{}
\end{figure}

@@

\section{Future Work}
On the theoretical side, we intend to investigate the semantic models of CLC. In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy. This has since been removed due to the additional difficulty of working with $Prop$ in the soundness proof of CILC. We believe that these difficulties are not intrinsic and $Prop$ could be added back with careful proof work.
